%!TEX root =  main.tex


\chapter{Access Management}

\section{Prelude}

In \emph{Information Technology} the topic \emph{Access Management}
describes endeavours of managing user access to data and system
features.\footnote{User access:  beside human users other \emph{actors}
do exist.  Automated processes, sensors, interfaces or batch jobs
interact the same way.  We use the term user as comprehensive form of
actor, which implies \emph{including} all forms of access.} The sibling
discipline \emph{Identity Management}, delivers input to Access
Management --- a perpetually changing data set, containing user
identities and roles. All decisions are finally based on this set.

Access Management shall distinguish between the \emph{good ones} and the
\emph{bad ones}.  The good ones (users allowed to access) shall be
enabled to access business transactions without undesired limitations in
terms of time, location or device used. The supplementary effort needed
to get access shall be minimized.  As a consequence productivity will
increase.

The other group, the bad ones, are those we do not want to access the
system. No matter what efforts they undertake to get access,
unauthorized use shall be prevented and misuse shall be identified and
reported.

Access Management delivers sustainability in those areas:

\begin{description}
    \item[Application Security] --- restrict users and programs, recognise
        system abuse, identify identity theft
    \item[User Productivity] --- providing users with ease access to their
        needs, exactly to their needs: not more, not less
    \item[Middleware technology] --- save investments with process
        centralization and software reuse, providing interfaces based
        on open internet protocols
    \item[Compliance] --- as additional value delivered, compliance itself
        usually won't work as impulse or driver
\end{description}




\section{In Real Life: Border Controls}

A good comparison to illustrate the main characteristics of
authentication is \emph{border control}. That one you'll most often find
on airports or on the actual boarder between countries.\footnote{There
    was a time you had to leave \emph{Schengen Area} to find this seldom
experience, nowadays there seems to be an inflation of borderline
checks.}

You leave one domain, or more important: you enter another domain. 
A domain that has a certain definition, borders, and authorities that 
execute policies (for good and for bad, however).  Those brave officers 
want you to identify yourself --- your job is to provide them data, exact 
and distinct data and some prove that you tell the truth.

The data is usually good enough to identify you among all human beings.
That quality of your prove must satisfy the procedure of those border
checks. With a valid passport you own a document from a common trusted
authority that covers this.

\begin{enumerate}
    \item identifying you among all others \emph{(identification)} and
    \item prove the correctness of your statement
        \emph{(authentication)}
\end{enumerate}

For you, as the traveler, the desired outcome of this procedure is your
valid entry without much delay.


\subsection{Alternative Flows}

Thinking more general, what other situations in or outcomes of this
situation are possible?

\begin{itemize}
    \item you could pretend to be another person, perhaps from another 
        country
    \item you could sneak in the country, preventing any check
    \item you could mix you passport by mistake in the cafeteria next 
        to the borderline.  You enter the country with a different identity 
        without knowing
    \item your passport is expired and you are not allowed to enter the 
        country
    \item etc.
\end{itemize}

All those situations have a direct mapping to use cases in the IT world.


\subsection{Not Authorized}

One more thing is in common: at border controls there is no check what
exactly you are allowed to do --- in the country or domain.

There might be a rough categorization implied during the authentication,
like people with a certain nationality would need a visa or people out
of a certain country are not allowed to enter at all.

But this is not the place where a work or residence permit would be
checked.\footnote{At least as far the author knows.}

Using our vocabulary, we are \emph{authenticated} but no authorization
check has been done.

